Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-04-09 paper/paper.tex
some polishing

\subsection{Partially Observable Markov Decision Processes}
It is our view that at a sufficiently coarse scale, the brain -- and therefore the DMN in particular --
is a physical system governed by the laws of
physics and can be locally described by Markov processes. Indeed, as noted in \cite{tegmark2016improved},
any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
step is sufficiently short, defining $s(t)$, the state at time $t$, as the position in state space.
The process has $memory$ if the next state depends not only on the current state but also on some finite
number of past states.
Rational probabilistic planning can be reformulated
as a standard memoryless Markov process by simply expanding the
definition of the state $s$ to include elements of the past.

@@

\subsection{Long-term rewards and action policies}
The behavior of the agent is governed by some kind of \textit{policy},
which maps states of the world to
a set of candidate actions to perform given this state. Starting a time $t=0$,
a policy $\pi$ generates a trajectory of action cascades as follows
\begin{eqnarray*}
  \begin{split}
    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect reward }R_0 = r(s_0, a_0)\\
    \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect reward }R_1 = r(s_1, a_1)\\
    \ldots\\
    \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect reward }R_{t} = r(s_{t}, a_{t})\\
    \ldots
  \end{split}
\end{eqnarray*}

@@

Since an action taken in the present moment $t$
might have repercussions long into the future, it turns out that the
quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
cumulative reward estimate which takes into account expected reward in the future.
A common approach to
modeling this accumulation is the time-discounted \textit{cumulative reward}% function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t = R_0 + \gamma R_1 + \gamma^2 R_2 + \ldots + \gamma^tR_t + \ldots
\end{equation}
This random variable\footnote{Random as it depends both on the environment's dynamic and the
  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
following a policy $\pi$.
\begin{mdframed}
  \textbf{Where is ``value'' buffered in the DMN ?
    % brain ?
    }
  The amygdala is known to be involved in significance evaluation while
  the NAc, which is connected to the vmPFC, is involved in reward evaluation.
\end{mdframed}
  The goal of the RL agent is then to adapt this policy so as to  maximize $G^\pi$. In
  the definition of cumulative reward $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
Setting $\gamma = 0$ corresponds to a perfectly shortsighted agent who is solely concerned about its immediate rewards.
Such an agent would have no horizon for the future,
which is not compatible with long-term goal planning
as potentially implemented in the DMN.
To allow a learning process to arise,
it is necessary that $0 < \gamma < 1$.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
% The reward discount factor
$\gamma$ can be seen as calibrating risk-seeking valency of the intelligent agent,
that is, the decision related to reward-delay trade-off schedules.
Such an agent puts relatively more emphasis on rewards expected in
a short future and pays less attention to
rewards expected at a later time, precisely rewards more than $1/(1 - \gamma)$ time steps into the future are disregarded.
This reduces the variance of expected rewards to be accumulated across action cascades by limiting the length of this cascades.
% given that there is more uncertainty in the farsighted future.
% It is also important to appreciate that a stochastic policy estimation
% is more advantageous in many RL settings.
% For instance, a deterministic strategy in playing rock-paper-scissors can
% be easily exploited and a uniformly random choice is optimal.

@@

\subsubsection{Efficient learning via value approximation and experience replay}
%As already mentioned in the previous section, Q-learning optimizes over the class of
% deterministic policies of the form \eqref{eq:qlearning}.
A successful RL theory needs ways to efficiently approximate the state-action value function
to allow for scaling to large state and action spaces.
In a state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
% (for example, $a$ = ``Read'', or ``Run'', or ``Laugh'', or ``Sympathize'', or ``Empathize'', etc.)
by sampling from its current policy, and collects a reward $r$,
% (measured on their internal currency of worldly values ?),
after which the environment transitions to a new state $s' \in \mathcal S$.
% \textbf{XXX Please try to smoothen the previous sentence}
At the end of this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced, which represents an exemplar
behavior of the agent and is recorded in replay memory buffer:
$D \leftarrow D \cup \{e\}$ (possibly discarding the oldest entries to make space).
Now, at iteration $k+1$, replay consists in sampling (uniform or importance-weighted
\footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
$(s, a, r, s') \sim \mathcal D$ from the replay memory $\mathcal D$.
The agent then tries to approximate
the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by the Bellman equation \eqref{eq:bellman}, namely
\begin{equation}
  y_k := r_k + \gamma \max_{a'} \tilde{Q}(s', a'|\theta_{k-1}),
\end{equation}
with the output of a parametrized regression model $(s,a)
\mapsto \tilde{Q}(s, a|\theta_{k-1})$. Here the function $(s,a) \mapsto Q(a,s|\theta_{k-1})$
% \textbf{XXX just double-checking, are the dots int he Q() correct?}
is an approximator for the $Q$-value operator,  parameterized by weights $\theta_{k}$
(synaptic strengths, etc.).

@@

Computing an optimal policy then corresponds to finding the parameters $\theta_{k}$ which minimize the following mean-squared loss function
\begin{equation}
  \mathcal L(\theta^Q_{k})
  % = \mathbb E_{(s, a, r, s') \sim \mathcal D}[(\tilde{Q}(s, a|\theta) - \tilde{Q}(s',a'|\theta^Q_{k}))^2]
  = \mathbb E_{(s, a, r, s') \sim \mathcal D}\left[\frac{1}{2}(\tilde{Q}(s, a|\theta_{k}) - y_k)^2\right].
  \label{eq:oracle}
\end{equation}
For instance, a general linear model with a kernel $\phi$ would be of the
form
$$\tilde{Q}(s, a|\theta) = \phi(s,a)^T\theta.$$
$\phi(s,a)$ would represent a high-level hand-crafted representation of the state-action pairs
$(s,a)$, as proposed by \citep{songNIPS2016}.
%  and Parr et al. 2003 (JMLR)].
A recently proposed, practically successful alternative approach
\citep{mnih2015,silver2016mastering} is to learn this
representation using a deep neural network, leading to the so-called \textit{Deep Q-learning} family of methods which
likely are state-of-the-art in RL.
The set of $\theta^Q$ parameters that instantiate the nonlinear interactions
between layers of the artifical neural network
may find a neurobiological correspondence in the adaptive strengths of axonal
connections between neurons from the different levels
of the neural processing hierarchy
\citep{mesulam1998sensation, taylor2015global}.

Learning
of the the entire model parameters can effectively be achieved
via \textit{backpropagation},
where reward prediction errors (RPES) -- also known as \textit{regret}-- percolate from lower to higher processing layers to modulate the choice of future actions
% online stochastic gradient descent is used to minimize the variational objective
% \eqref{eq:oracle}.

@@

\subsection{Predictive coding hypothesis}
Predictive coding mechanisms
\citep{clark2013whatever, friston2008hierarchical}
are a frequently evoked idea in the context of default mode function
\citep{bar2007}.
Cortical responses are explained as
emerging from continuous functional interaction between
higher and lower levels of the neural processing hierarchy.
Feed-forward sensory processing is constantly calibrated by
top-down modulation from more multi-modal and associative brain regions
further away from primary sensory cortical regions.
The dynamic interplay between cortical processing levels
may enable inferences about the world by reconciling
gaps between fresh sensory input and expectations computed
from stored prior information.
At each stage of neural processing,
an internally generated prediction of aspects of sensory sensation is
directly compared against the actual environmental input.
A prediction error at one of the processing levels
incurs plasticity changes of neuronal back projections
(i.e., adapting model parameters)
to allow for gradually improved future prediction of the environment.
In this way,
the predictive coding hypothesis offers explanations for
the constructive, generative nature of sensory perception
\citep{friston2010free} and
the intimate relation of motor action to sensory expectations
\citep{wolpert1995internal, kording2004bayesian}.
Contextual sensorimotor integration of perception-action cycles
may be maintained by top-down modulation
by a-priori information about the environment.

Predictive coding processes
permit updates of the internal representation of the environment
to best accommodate and prepare the organism
for the constant influx of sensory examples and
for reacting to the environment.
There are hence a number of common properties
between the predictive coding account
and the proposed formal account.
Importantly,
a generative model of how perceived sensory cues arise in the
world would be incorporated into
the current neuronal wiring.
Further,
both frameworks are plausibilized by neuroscientific evidence that suggest
the human brain to be a ``statistical organ'' with the biological purpose to
generalize from the past to new experiences.
Neuroanatomically, axonal
back projections probably outnumber by far the axonal input projections
existing in the primate brain \citep{salin1995corticocortical}.
Moreover,
both accounts provide a parsimonious explanation why the
human brain decreases processing load devoted to incoming information
when the environment is predictable because the internal generative
model only requires updating following descrepancies between
environmental reality and its internal reinstantiation.
Increased computation resources are however allocated
when novel stimuli or
unexpected events are encountered by the organism.
The predictive coding and formal account
naturally evoke a mechanism of brain plasticity in that
neuronal wiring gets increasingly adapted
when faced by unprecedented environmental challenges.

While sensory experience is a generative process from both views,
the predictive coding account frames
sensory perception of the external world
as a generative experience due to the modulatory top-down influence at
various hierarchical levels of sensory input processing.
In our RL view, instead,
the Markov property of our sequential decision making framework entails
the environmental model to be
\textbf{effectively expressed in the current state of the human agent
XXX guys, please double-check}.
Further,
the hierarchical processing aspect from predictive coding
is re-expressed in our account in form of
nested prediction of probable upcoming actions, states, and rewards.
Both frameworks model the consequences of action.
The predictive coding framework is typically explained without
explicit parameterization of the agent's horizon and
has a tendency to be presented as emphasizing prediction about the
immediate future.
In the present account, the horizon of that
look into the future is explicitly expressed in the $\gamma$ parameter
of the Bellman equation.
% \item \suggestremove{Mismatch negativity of PC is immanent in reward prediction error
% by the difference between actually predicted reward of an action given
% a state and the reward predicted by non-linear regression with graded
% discounting of future rewards.}
Finally,
the process of adapting the neuronal connections
for improved top-down modulation
takes the concrete form of gradient computation and
back-propagation in our RL implementation.
It is however important to note that
the neurobiological plausibility of
the back-propagation procedure is controversal
\citep{goodfellow2016deep}.
In sum,
the formal account may serve as a concrete implementation of
the long-standing predictive coding account from cognitive neuroscience.
MPDs have the advantage of exposing an explicit
mechanism of the horizon of future considerations or
how the internal model of the world gets updated,
and
more explicitly reconciles sensory input processing and action consideration.
%
Recasting DMN function in terms of MDPs therefore naturally incorporate
a majority of aspects of the prediction coding hypothesis.

\subsection{Semantic hypothesis}
Another frequently proposed cognitive account of DMN function revolves
around forming logical associations and analogies between
the current experience and
the conceptual knowledge derived from past experiences
\citep{bar2007proactive, binder1999conceptual}.
Analogies might naturally tie incoming novel sensory stimuli to
explicit world knowledge (i.e., semantics) extracted from the environment
\citep{bar2009proactive}.
The encoding of complex environments could thus be facilitated
by relative similarity associations between states.
%
Going beyond human language itself,
semantic building blocks may provide the basis for
mentally envisioning non-existent scenarios
that would optimize behavior in the environment
by constantly simulating possible future events.
Such cognitive processes can afford
the internal generation of necessary information
that is not presented in the surrounding environment
by recombining building blocks of
concept knowledge and episodic memories
\citep{hassabis2009construction}.
Indeed, in aging humans, remembering the past and imaging the future
equally decreased in the level of detail and were associated with
concurrent deficits in forming and integrating relationships between
items \citep{addis2008age, spreng2006temporal}.
Further,
a constructive account can explain the reciprocal relationship
between an egocentric first person perspective and
an allocentric bird’s eye perspective immersed in
self-reflection, semantic associations, and autobiographical memories.
%
Cognitive aspects of egocentric-allocentric switching
are also closely related to episodic memory, language, problem solving,
planning, estimating other people's thoughts, and spatial navigation.
These neural processes are likely to
build on abstract world knowledge and abstract assocations
for binding the constituent elements in mental scene construction
\citep{schacter2007remembering}.
Such scene generation processes could contribute to interpreting the
present and foretelling the future.
Mental scene construction has been proposed
to imply a distinction between
environmental perception and internally generated mind-wandering
\citep{buckner2007self}.
These authors stated that
``A computational model [...] will probably require a form of
regulation by which perception of the current world is suppressed
while simulation of possible alternatives are constructed,
followed by a return to perception of the present.''.
The semantic hypothesis is for instance supported by evidence in animals that
could learn a \textit{cognitive map of the environment},
even without reward incentives, and exploit it later
for other means \citep{tolman1948cognitive}.

In comparison,
both the semantic hypothesis and MDPs expose mechanisms of
how action considerations could be mentally explored.
In both accounts,
there is also no evidence to indicate that predictions of various
levels of complexity, abstraction, timescale and purpose
use mechanisms that are qualitatively different. This concurs with
DMN activity increases across time, space, and content domains
demonstrated in many neuroimaging studies \citep{spreng2009common}.
The semantic hypothesis
and MDP account provide explanations why hippocampus lesion does
not only impair retrieving memories, but also hypothetical and future
thinking \citep{hassabis2007patients}.
While both semantic hypothesis and
our formal account propose memory-based internally
generated information for probabilistic mental models of action outcomes,
MDPs render explicit the grounds on which the final action is
eventually chosen, namely, the estimated cumulative reward.
\textbf{The notion of semantic knowledge associations is
incorporated into the MDP as the Markov property,
that is, the current state directly results from the
agent's history of states and actions. The learned
value matrix and action transition schedules drive
stimulus processing and action choice in the present.}
The MDPs naturally integrate the former egocentric
(more related to current action, state and reward) and the later
allocentric (more related to past and future actions, states, and rewards)
angles on the world in a same optimization problem.
Further,
the semantic account of DMN function does not offer
a mechanistic explanation how explicit world knowledge and semantic analogies thereof
lead to prediction of future actions and states.
The semantic hypothesis does not explain why memory recall
for scene construction in humans is typically fragmentary and noisy
instead of accurate and reliable.
In contrast to existing accounts on semantics and
mental scene construction, the random and creative aspects of DMN function
are explained in MDPs by the advantages of stochastic optimization.
Yet, the MDP framework provides an algorithmic explanation in that
stochasticity of the parameter space explored
by the Monte Carlo solvers achieves better fine-tuning of the
action policies and estimation of expected reward outcomes.
That is, the purposeful stochasticity of policy and value estimation
in MDPs provides a candidate explanation for why humans
have evolved imperfect noisy memories
as the more advantageous adaptation.
In sum, episodic scene construction according to the semantic
hypothesis is lacking an explicit time and incentive model;
neither does it explain how randomness of human mental experience
can be beneficial.

>>>>

\subsection{Partially Observable Markov Decision Processes}
It is our view that at a sufficiently coarse scale, the brain --and therefore the DMN in particular--
is a physical system governed by the laws of
physics and can be locally described by Markov processes. Indeed, as noted in \cite{tegmark2016improved},
any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
step is sufficiently short
% , defining $s(t)$, the state at time $t$, as the position in state space.
The process has $memory$ if the next state depends not only on the current state but also on some finite
number of past states.
Rational probabilistic planning can be reformulated
as a standard memoryless Markov process by simply expanding the
definition of the state $s$ to include elements of the past.

@@

\subsection{Long-term rewards and action policies}
The behavior of the agent is governed by some kind of \textit{policy},
which maps states of the world to
a set of candidate actions to perform given this state. Starting a time $t=0$,
a policy $\pi$ generates a trajectory of action cascades / trajectories as follows
\begin{eqnarray*}
  \begin{split}
    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect reward }R_0 = r(s_0, a_0)\\
    \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect reward }R_1 = r(s_1, a_1)\\
    \ldots\\
    \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect reward }R_{t} = r(s_{t}, a_{t})\\
    \ldots
  \end{split}
\end{eqnarray*}

@@

Since an action taken in the present moment $t$
might have repercussions long into the future, it turns out that the
quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
cumulative reward estimate which takes into account expected reward in the future.
A common approach to
modeling this accumulation is the time-discounted \textit{cumulative reward}% function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t = R_0 + \gamma R_1 + \gamma^2 R_2 + \ldots + \gamma^tR_t + \ldots
\end{equation}
This random variable\footnote{Random as it depends both on the environment's dynamic and the
  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
following a policy $\pi$.
\begin{mdframed}
  \textbf{Where is ``value'' buffered in the DMN ?
    % brain ?
    }
  The amygdala is known to be involved in significance evaluation while
  the NAc, which is connected to the vmPFC, is involved in reward evaluation.
\end{mdframed}
  The goal of the RL agent is then to adapt this policy so as to  maximize $G^\pi$. In
  the definition of cumulative reward $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
Setting $\gamma = 0$ corresponds to a perfectly shortsighted agent who is solely concerned about its immediate rewards.
Such an agent would have no horizon for the future,
which is not compatible with long-term goal planning
as potentially implemented in the DMN.
To allow a learning process to arise,
it is necessary that $0 < \gamma < 1$.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
% The reward discount factor
$\gamma$ can be seen as calibrating risk-seeking valency of the intelligent agent,
that is, the decision related to reward-delay trade-off schedules.
Such an agent puts relatively more emphasis on rewards expected in
a short future and pays less attention to
rewards expected at a later time, precisely rewards more than $1/(1 - \gamma)$ time steps into the future are disregarded.
This reduces the variance of expected rewards accumulated across action cascades by limiting the length of this cascades.
% given that there is more uncertainty in the farsighted future.
% It is also important to appreciate that a stochastic policy estimation
% is more advantageous in many RL settings.
% For instance, a deterministic strategy in playing rock-paper-scissors can
% be easily exploited and a uniformly random choice is optimal.

@@

\subsubsection{Efficient learning via value approximation and experience replay}
%As already mentioned in the previous section, Q-learning optimizes over the class of
% deterministic policies of the form \eqref{eq:qlearning}.
A successful RL theory needs ways to efficiently approximate the state-action value function
to allow for scaling to large state and action spaces.
In a state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
% (for example, $a$ = ``Read'', or ``Run'', or ``Laugh'', or ``Sympathize'', or ``Empathize'', etc.)
by sampling from its current policy, and collects a reward $r$,
% (measured on their internal currency of worldly values ?),
after which the environment transitions to a new state $s' \in \mathcal S$.
% \textbf{XXX Please try to smoothen the previous sentence}
At the end of this cycle, a new \textit{experience} $e = (s,a,r,s')$ is produced, which represents an exemplar
behavior of the agent and is recorded in replay memory buffer
(possibly discarding the oldest entries to make space):
$\mathcal D \leftarrow \text{append}(\mathcal D, e)$.
At iteration $k+1$, replay then consists in sampling (uniform or importance-weighted
\footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
$(s, a, r, s') \sim \mathcal D$ from the replay memory $\mathcal D$.
The agent then tries to approximate
the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by the Bellman equation \eqref{eq:bellman}, namely
\begin{equation}
  y_k := r_k + \gamma \max_{a'} \tilde{Q}(s', a'|\theta_{k-1}),
\end{equation}
with the output of a parametrized regression model $(s,a)
\mapsto \tilde{Q}(s, a|\theta_{k-1})$. Here the function $(s,a) \mapsto \tilde{Q}(a,s|\theta_{k-1})$
% \textbf{XXX just double-checking, are the dots int he Q() correct?}
is an approximator for the $Q$-value operator,  parameterized by weights $\theta_{k}$
(synaptic strengths, etc.).

@@

Computing an optimal policy then corresponds to finding the parameters $\theta_{k}$ which minimize the following mean-squared loss function
\begin{equation}
  \mathcal L(\theta^Q_{k})
  % = \mathbb E_{(s, a, r, s') \sim \mathcal D}[(\tilde{Q}(s, a|\theta) - \tilde{Q}(s',a'|\theta^Q_{k}))^2]
  = \mathbb E_{(s, a, r, s') \sim \mathcal D}\left[\frac{1}{2}(\tilde{Q}(s, a|\theta_{k}) - y_k)^2\right].
  \label{eq:oracle}
\end{equation}
For instance, a general linear model with a kernel $\phi$ would be of the
form
$$\tilde{Q}(s, a|\theta) = \phi(s,a)^T\theta.$$
$\phi(s,a)$ would represent a high-level hand-crafted representation of the state-action pairs
$(s,a)$, as proposed by \citep{songNIPS2016}.
%  and Parr et al. 2003 (JMLR)].
A recently proposed, practically successful alternative approach
\citep{mnih2015,silver2016mastering} is to learn this
representation using a deep neural network, leading to the so-called \textit{Deep Q-learning} family of methods which
likely are state-of-the-art in RL.
The set of $\theta$ parameters that instantiate the nonlinear interactions
between layers of the artifical neural network
may find a neurobiological correspondence in the adaptive strengths of axonal
connections between neurons from the different levels
of the neural processing hierarchy
\citep{mesulam1998sensation, taylor2015global}.

Learning
of the the entire model parameters can effectively be achieved
via \textit{backpropagation},
where reward prediction errors (RPEs) -- also known as \textit{regret}-- percolate
from lower to higher processing layers to modulate the choice of future actions
% online stochastic gradient descent is used to minimize the variational objective
% \eqref{eq:oracle}.

@@

\subsection{Predictive coding hypothesis}
Predictive coding mechanisms
\citep{clark2013whatever, friston2008hierarchical}
are a frequently evoked idea in the context of default mode function
\citep{bar2007}.
According to this view,
cortical responses emerge from continuous functional interaction between
higher and lower levels of the neural processing hierarchy.
Feed-forward sensory experience is constantly calibrated by
top-down modulation realized by more multi-modal and associative brain regions
further away from the primary sensory cortices.
The dynamic interplay between cortical processing levels
may enable inferences about the world by reconciling
gaps between fresh sensory input and expectations computed
from stored prior information.
At each stage of neural processing,
an internally generated prediction of aspects of sensory input is
directly compared against the actual environmental input.
A prediction error at one of the processing levels,
i.e., the difference between sensory input
and internally predicted sensation,
incurs plasticity changes of neuronal back projections
(i.e., adapting model parameters)
to allow for gradually improved future prediction of the environment.
In fact, axonal
back projections probably outnumber by far the axonal input projections
existing in the primate brain \citep{salin1995corticocortical}.
This can explain the constructive, generative nature of sensory perception
\citep{friston2010free} and
why motor action is probably intimately related to sensory expectations
\citep{wolpert1995internal, kording2004bayesian}.
Contextual sensorimotor integration of perception-action cycles
may hence be maintained by top-down modulation
by a-priori information about the environment.

This process permits updates of the internal representation of the environment
to best accommodate and prepare for the constant influx of sensory examples.
There are hence a number of common properties
between the predictive coding account
and the proposed formal account.
Importantly,
the generative model of how perceived sensory cues arise in the
world would be incorporated into
the current neuronal wiring.
Further,
both frameworks are plausibilized by neurobiological evidence that suggest
the human brain to be a ``statistical organ'' with the biological purpose to
generalize from the past to new experiences.
Both accounts also provide a parsimonious explanation why the
human brain decreases processing load devoted to incoming information
when the environment is predictable and allocates increased
computation resources when novel stimuli or unexpected events are encountered.
The predictive coding and formal account
naturally evoke a mechanisms of brain plasticity in that
neuronal wiring gets increasingly adapted
when faced by unprecedented environmental
challenges.

While sensory experience is a generative process from both views,
the predictive coding account frames
sensory perception of the external world
as a generative experience due to the modulatory top-down influence at
various hierarchical levels of sensory input processing. Instead,
the Markov property of our sequential decision making framework.
In our RL view, the environmental model is
\textbf{effectively expressed in the current state of the human agent
XXX guys, please double-check}.
The hierarchical processing aspect from predictive coding
is re-expressed in our account in form of
nested prediction of probable upcoming actions, states, and rewards.
Both frameworks model the consequences of action.
The predictive coding framework is typically explained without
explicitly parameterization of the agent's horizon and
has a tendency to be presented as emphasizing prediction about the
immediate future.
In the present account, the horizon of that
look into the future is explicitly expressed in the $\gamma$ parameter
of the Bellman equation.
% \item \suggestremove{Mismatch negativity of PC is immanent in reward prediction error
% by the difference between actually predicted reward of an action given
% a state and the reward predicted by non-linear regression with graded
% discounting of future rewards.}
The process of adapting the neuronal connections
for improved top-down modulation
takes the concrete form of gradient computation and
back-propagation in our RL implementation.
It is however important to note that
the neurobiological plausibility of
the back-propagation procedure is controversal
\citep{goodfellow2016deep}.
In sum,
the formal account may serve as a concrete implementation of
the predictive coding account
from cognitive neuroscience.
MPDs have the advantage of exposing an explicit
mechanism of the horizon of future considerations or
how the internal model of the world gets updated,
and
more explicitly reconciles sensory input processing and action output preparation.
%
As such, recasting DMN function in terms of MDPs naturally incorporates
a majority of aspects of the prediction coding hypothesis.

\subsection{Semantic hypothesis}
Another frequently proposed cognitive account of DMN function revolves
around forming logical associations and analogies between
the current experience and
the conceptual knowledge derived from past experiences
\citep{bar2007proactive, binder1999conceptual}.
Analogies might naturally tie incoming novel sensory stimuli to
explicit world knowledge (i.e., semantics) extractable from the environment
\citep{bar2009proactive}.
In this way, the encoding of complex environments could be facilitated
by relative similarity associations between states.
%
Going much beyond human language ifself,
semantic building blocks provide the basis to
mentally envision non-existent scenarios
that would improve optimal behavior in the environment
by simulating future events.
Such cognitive processes can afford
the internal generation of necessary information
that is not presented in the surrounding environment
by recombining building blocks of
concept knowledge and episodic memories
\citep{hassabis2009construction}.
Indeed, in aging humans, remembering the past and imaging the future
equally decreased in the level of detail and are associated with
concurrent deficits in forming and integrating relationships between
items \citep{addis2008age, spreng2006temporal}.
Further,
a constructive account explains the reciprocal relationship
between an egocentric first person perspective and
an allocentric bird’s eye perspective immersed in
self-reflection, social knowledge, and autobiographical memories.
%
Cognitive aspects of egocentric-allocentric switching
are also closely related to episodic memory, language, problem solving,
planning, estimating other people's thoughts, and spatial navigation
as these necessitate abstract world knowledge and abstract assocations
for binding the constituent elements in mental scene construction
\citep{schacter2007remembering}.
These scene generation processes could contribute to interpreting the
present and foretelling the future.
This view is for instance supported by evidence in animals that
could learn a \textit{cognitive map of the environment},
even without reward incentives, and exploit it later
for other means \citep{tolman1948cognitive}.

\begin{itemize}
  \item Both the semantic hypothesis and MDPs expose mechanisms of
  how alternative decision trees could be mentally explored.
  \item In both semantic hypothesis and MDPs,
  there is no evidence to indicate that predictions of various
  levels of complexity, abstraction, timescale and purpose
  use mechanisms that are qualitatively different. This concurs with
  DMN activity increases across time, space, and content domains
  demonstrated in many neuroimaging studies. The semantic hypothesis
  and RL account provide explanations why hippocampus lesion does
  not only impair retrieving memories, but also hypothetical and future
  thinking \citep{hassabis2007patients}.
  \item The notion of semantic or knowledge association is
  incorporated into the MDP as the Markov property,
  that is, the current state directly results from the
  agent's history of states and actions. The learned
  value matrix and action transition schedules drive
  stimulus processing and action choice in the present.
  \item Mental scene construction has been proposed by some authors
  \citep{buckner2007self} to imply a distinction between
    environmental perception and internally generated mind-wandering\footnote{
      \textbf{XXX Elvis: This is more than just a footnote. Make this part of the paper
      and factorize the lot. Also you probably need a citation here!!!}
  ``A computational model of how such a process might be structured
  is far from being defined, but it will probably require a form of
  regulation by which perception of the current world is suppressed
  while simulation of possible alternatives are constructed,
  followed by a return to perception of the present.''\citep{buckner2007self}}.
  The MDPs naturally integrate the former egocentric
  (more related to current action, state and reward) and the later
  allocentric (more related to past and future actions, states, and rewards)
  angles on the world in a same optimization problem.
  \item The semantic account of DMN function does not offer
  a mechanistic explanation how explicit world knowledge and semantic analogies thereof
  lead to prediction of future actions and states.
  \item In contrast to existing accounts on semantics and
  mental scene construction, the random and creative aspects of DMN function
  are explained in MDPs by the advantages of stochastic optimization.
  \item The semantic hypothesis does not explain why memory recall
  for scene construction in humans is typically fragmentary and noisy
  instead of accurate and reliable.
  Yet, the MDP framework provides an algorithmic explanation in that
  stochasticity of the parameter space explored
  by the Monte Carlo solvers achieves better fine-tuning of the
  action policies and estimation of expected reward outcomes.
  That is, the purposeful stochasticity of policy and value estimation
  in MDPs provides a candidate explanation for why humans
  have evolved imperfect noisy memories
  as the more advantageous adaptation.
  \item While both semantic and MDP account propose memory-based internally
  generated information for probabilistic mental models of action outcomes,
  only MDPs render explicit the grounds on which the final action is
  eventually chosen (i.e., the estimated cumulative reward).
  \item In sum, episodic scene construction according to the semantic
  hypothesis is lacking an explicit time and incentive model;
  neither does it explain how randomness of human mental experience
  can be beneficial.
\end{itemize}
